Cluster Analysis: Identifying Parkinson's Disease Subtypes

### 2015-07-22 writeup/abbreviated/abbreviated.tex
Update abbr. writeup with new data, corrplot

Combination of non-motor scale (NMS) symptoms and standard motor symptoms.

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
    Name & Type & Format & Description \\
    \hline
    nms\_d1 & byte & \%8.0g & cardiovascular \\
    nms\_d2 & byte & \%8.0g & sleep/fatigue \\
    nms\_d3 & byte & \%8.0g & mood/cognition \\
    nms\_d4 & byte & \%8.0g & percep/hallucinations \\
    nms\_d5 & byte & \%8.0g & attention/memory \\
    nms\_d6 & byte & \%8.0g & gastrointestinal \\
    nms\_d7 & byte & \%8.0g & urinary \\
    nms\_d8 & byte & \%8.0g & sexual function \\
    nms\_d9 & byte & \%8.0g & miscellaneous \\
    tremor & float & \%9.0g & tremor \\
    bradykin & float & \%9.0g & bradykinesia\tablefootnote{Impaired ability to
    adjust the body's position.} \\
    rigidity & float & \%9.0g & rigidity \\
    axial & float & \%9.0g & axial\tablefootnote{Issues affecting the middle of
    the body.} \\
    pigd & float & \%9.0g & postural instability and gait difficulty \\
  \end{tabular}
  \caption{Selected Features and Details}
  \label{tab:selected-features}
\end{table}

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
  Name  &       $\mu$ & $\sigma$ & min-max \\
         \hline
nms\_d1&   1.73&  3.35&   0-24 \\
nms\_d2&   8.75&  8.70&   0-48 \\
nms\_d3&   8.68& 11.55&   0-60 \\
nms\_d4&   1.64&  3.86&   0-33 \\
nms\_d5&   5.42&  7.43&   0-36 \\
nms\_d6&   5.53&  6.79&   0-36 \\
nms\_d7&   8.08&  8.94&   0-36 \\
nms\_d8&   3.52&  5.97&   0-24 \\
nms\_d9&   7.13&  7.79&   0-48 \\
tremor&   2.59&  2.58&   0-12 \\
bradykin& 2.40&  1.41&   0-6 \\
rigidity& 2.24&  1.36&   0-6 \\
axial&    3.25&  2.68&   0-12 \\
pigd&     3.31&  2.71&   0-12 \\
  \end{tabular}
  \caption{Descriptive Statistics}
  \label{tab:descriptive-statistics}
\end{table}

@@

\begin{table}[h]
  \centering
  \caption{Cluster statistics}
  \label{tab:cluster_stats}
  \begin{tabular}{l|l}
    Cluster & $n$ \\
    \hline
    1 & 79 \\
    2 & 394 \\
    3 & 275 \\
    4 & 153 \\
  \end{tabular}
\end{table}

\subsection{Decision tree}
\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l|l|l}
    $k$ & CP\tablefootnote{Complexity Parameter} & CV Xerror\tablefootnote{10-fold cross
    validation} & Root Feature &
    Root Error & Figure \\
    \hline
    4 & 0.0100 & 0.255 & pigd $<$ 2.5 & 0.563 & Figure~\ref{fig:kmeans-dtree-4} \\
  \end{tabular}
  \caption{$k$-kmeans decision trees statistics}
  \label{tab:k-means-dtrees}
\end{table}

@@

Post-hoc analysis using Tukey's HSD:
\begin{verbatim}
age insignificant differences:
         diff         lwr        upr        p adj
4-3 -2.271990  -4.7160927  0.1721117 7.920343e-02
sex insignificant differences:
           diff        lwr        upr     p adj
2-1 -0.09275204 -0.2454183 0.05991417 0.4000134
3-1 -0.14660529 -0.3046921 0.01148156 0.0803067
4-1  0.04757177 -0.1240051 0.21914864 0.8917054
3-2 -0.05385325 -0.1511666 0.04346014 0.4843788
pdonset insignificant differences:
           diff       lwr      upr     p adj
2-1 -0.90162565 -4.302019 2.498768 0.9037828
3-1 -0.05040276 -3.571532 3.470727 0.9999820
4-1 -0.53727145 -4.358869 3.284326 0.9837825
3-2  0.85122289 -1.316276 3.018722 0.7431486
4-2  0.36435420 -2.263251 2.991959 0.9844187
4-3 -0.48686869 -3.268952 2.295214 0.9695444
nms_d1 insignificant differences:
          diff         lwr       upr        p adj
3-2  0.5565021 -0.01586385  1.128868 6.018674e-02
nms_d3 insignificant differences:
          diff        lwr        upr       p adj
4-1  -1.192024  -4.493510   2.109461 0.789133802
nms_d4 insignificant differences:
          diff        lwr       upr        p adj
3-2  0.2951915 -0.3318339  0.922217 6.195406e-01
nms_d5 insignificant differences:
           diff         lwr       upr        p adj
3-2   0.7841071  -0.4814825  2.049697 0.3821565990
nms_d8 insignificant differences:
          diff       lwr      upr     p adj
4-1 -0.9010507 -2.846119 1.044017 0.6318563
3-2  0.7255838 -0.377602 1.828770 0.3280035
nms_d9 insignificant differences:
          diff         lwr       upr       p adj
4-1  0.7048896  -1.6510530  3.060832 0.867951568
tremor insignificant differences:
          diff       lwr        upr      p adj
3-1 -0.2662831 -1.064520 0.53195417 0.82613187
4-2 -0.5667198 -1.162396 0.02895624 0.06895758
rigidity insignificant differences:
          diff         lwr        upr        p adj
4-2  0.2310142 -0.03536748  0.4973959 1.153952e-01
\end{verbatim}

@@

\begin{table}[h]
  \centering
  \caption{Features ranked by information gain}
  \label{tab:info_gain}
  \begin{tabular}{l|l}
    variable & information gain \\
    \hline
    axial      & 0.20640691 \\
    cisitot      & 0.20008571 \\
    pigd      & 0.18193982 \\
    nms\_d2      & 0.13178572 \\
    nms\_d9      & 0.12116024 \\
    bradykin      & 0.11966097 \\
    nms\_d3      & 0.09421859 \\
    rigidity      & 0.09260628 \\
    nms\_d5      & 0.07579997 \\
    nms\_d4      & 0.07438784 \\
    nms\_d6      & 0.06620599 \\
    nms\_d7      & 0.05574956 \\
    nms\_d1      & 0.05509838 \\
    tremor      & 0.04140473 \\
    nms\_d8      & 0.03786173 \\
    durat\_pd      & 0.02794420 \\
    age      & 0.00000000 \\
    sex      & 0.00000000 \\
    pdonset      & 0.00000000 \\
  \end{tabular}
\end{table}

@@

\subsubsection{Bradykinesia and rigidity}

Figures~\ref{fig:bradykin-cisitot} and~\ref{fig:rigidity-cisitot}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{bradykin-cisitot.pdf}
  \caption{Relationship between bradykinesia and cisitot}
  \label{fig:bradykin-cisitot}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{bradykin-cisitot-condensed.pdf}
  \caption{Relationship between bradykinesia and cisitot (condensed)}
  \label{fig:bradykin-cisitot-condensed}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{bradykin-cisitot-all.pdf}
  \caption{Relationship between bradykinesia and cisitot (entire data)}
  \label{fig:bradykin-cisitot-all}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{rigidity-cisitot.pdf}
  \caption{Similar relationship between rigidity and cisitot}
  \label{fig:rigidity-cisitot}
\end{figure}

\section{Other Work}

>>>>

Combination of non-motor scale (NMS) symptoms and standard motor symptoms.
Note: PIGD was deleted after 2015-07-16 meeting.

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
    Name & Type & Format & Description \\
    \hline
    nms\_d1 & byte & \%8.0g & cardiovascular \\
    nms\_d2 & byte & \%8.0g & sleep/fatigue \\
    nms\_d3 & byte & \%8.0g & mood/cognition \\
    nms\_d4 & byte & \%8.0g & percep/hallucinations \\
    nms\_d5 & byte & \%8.0g & attention/memory \\
    nms\_d6 & byte & \%8.0g & gastrointestinal \\
    nms\_d7 & byte & \%8.0g & urinary \\
    nms\_d8 & byte & \%8.0g & sexual function \\
    nms\_d9 & byte & \%8.0g & miscellaneous \\
    tremor & float & \%9.0g & tremor \\
    bradykin & float & \%9.0g & bradykinesia\tablefootnote{Impaired ability to
    adjust the body's position.} \\
    rigidity & float & \%9.0g & rigidity \\
    axial & float & \%9.0g & axial\tablefootnote{Issues affecting the middle of
    the body.} \\
  \end{tabular}
  \caption{Selected Features and Details}
  \label{tab:selected-features}
\end{table}

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
  Name  &       $\mu$ & $\sigma$ & min-max \\
         \hline
nms\_d1&   1.73&  3.35&   0-24 \\
nms\_d2&   8.75&  8.70&   0-48 \\
nms\_d3&   8.68& 11.55&   0-60 \\
nms\_d4&   1.64&  3.86&   0-33 \\
nms\_d5&   5.42&  7.43&   0-36 \\
nms\_d6&   5.53&  6.79&   0-36 \\
nms\_d7&   8.08&  8.94&   0-36 \\
nms\_d8&   3.52&  5.97&   0-24 \\
nms\_d9&   7.13&  7.79&   0-48 \\
tremor&   2.59&  2.58&   0-12 \\
bradykin& 2.40&  1.41&   0-6 \\
rigidity& 2.24&  1.36&   0-6 \\
axial&    3.25&  2.68&   0-12 \\
  \end{tabular}
  \caption{Descriptive Statistics}
  \label{tab:descriptive-statistics}
\end{table}

@@

\begin{table}[h]
  \centering
  \caption{Cluster statistics}
  \label{tab:cluster_stats}
  \begin{tabular}{l|l}
    Cluster & $n$ \\
    \hline
    1 & 189 \\
    2 & 88 \\
    3 & 221 \\
    4 & 406 \\
  \end{tabular}
\end{table}

\subsection{Decision tree}
% \begin{table}[h]
  % \centering
  % \begin{tabular}{l|l|l|l|l|l}
    % $k$ & CP\tablefootnote{Complexity Parameter} & CV Xerror\tablefootnote{10-fold cross
    % validation} & Root Feature &
    % Root Error & Figure \\
    % \hline
    % 4 & 0.0100 & 0.255 & pigd $<$ 2.5 & 0.563 & Figure~\ref{fig:kmeans-dtree-4} \\
  % \end{tabular}
  % \caption{$k$-kmeans decision trees statistics}
  % \label{tab:k-means-dtrees}
% \end{table}

@@

Post-hoc analysis using Tukey's HSD:
\begin{verbatim}
age insignificant differences:
          diff       lwr       upr     p adj
3-1  0.9947808 -1.458184 3.4477455 0.7236845
4-1 -1.2838898 -3.464063 0.8962832 0.4284274
sex insignificant differences:
           diff        lwr        upr      p adj
2-1 -0.05044493 -0.2106093 0.10971941 0.84945412
4-1 -0.09897829 -0.2082638 0.01030726 0.09181043
3-2 -0.12633690 -0.2827741 0.03010026 0.16087872
4-2 -0.04853336 -0.1944676 0.09740091 0.82744866
4-3  0.07780354 -0.0259428 0.18154987 0.21607772
pdonset insignificant differences:
          diff        lwr      upr     p adj
2-1  2.9315777 -0.6232172 6.486373 0.1466742
3-1  1.7136632 -1.0153886 4.442715 0.3699280
4-1  0.7453932 -1.6801637 3.170950 0.8585776
3-2 -1.2179144 -4.6899860 2.254157 0.8033301
4-2 -2.1861845 -5.4251477 1.052779 0.3049434
4-3 -0.9682701 -3.2708860 1.334346 0.7004488
durat_pd insignificant differences:
          diff       lwr        upr        p adj
3-1 -0.7188824 -2.140915  0.7031499 5.624040e-01
cisitot insignificant differences:
          diff         lwr       upr        p adj
3-1  0.4942421  -0.4388731  1.427357 5.228228e-01
nms_d1 insignificant differences:
          diff        lwr        upr        p adj
4-3 -0.3798787 -0.9604053  0.2006478 3.325894e-01
nms_d4 insignificant differences:
          diff        lwr        upr        p adj
4-3 -0.3362459 -0.9972012  0.3247094 5.571480e-01
nms_d5 insignificant differences:
           diff        lwr         upr        p adj
4-3  -0.4117201  -1.743630   0.9201902 8.564409e-01
nms_d8 insignificant differences:
          diff        lwr        upr        p adj
4-3 -0.9953302 -2.1560641  0.1654036 1.220509e-01
nms_d9 insignificant differences:
          diff       lwr        upr      p adj
2-1  0.8708514 -1.297413 3.03911557 0.72966265
4-3 -1.3221920 -2.726684 0.08229967 0.07350641
tremor insignificant differences:
          diff         lwr        upr        p adj
4-1  0.3346105 -0.19270261  0.8619236 3.603863e-01
\end{verbatim}

@@

\begin{table}[h]
  \centering
  \caption{Features ranked by information gain}
  \label{tab:info_gain}
  \begin{tabular}{l|l}
    variable & information gain \\
    \hline
    bradykin      & 0.31574672 \\
    rigidity      & 0.29560018 \\
    nms\_d2      & 0.24218407 \\
    cisitot      & 0.22920103 \\
    axial      & 0.22780750 \\
    nms\_d3      & 0.20480570 \\
    nms\_d9      & 0.15782743 \\
    nms\_d7      & 0.15290569 \\
    nms\_d5      & 0.14454931 \\
    nms\_d6      & 0.14025139 \\
    nms\_d1      & 0.13212756 \\
    tremor      & 0.10937168 \\
    nms\_d4      & 0.10710526 \\
    nms\_d8      & 0.10005480 \\
    durat\_pd      & 0.02876190 \\
    age      & 0.02346158 \\
    sex      & 0.00000000 \\
    pdonset      & 0.00000000 \\
  \end{tabular}
\end{table}

@@

\subsubsection{One vs all decision trees}

Figures~\ref{fig:1va},~\ref{fig:2va},~\ref{fig:3va},~\ref{fig:4va}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{dtree-1va-pruned.pdf}
  \caption{Cluster 1 vs all}
  \label{fig:1va}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{dtree-2va-pruned.pdf}
  \caption{Cluster 2 vs all}
  \label{fig:2va}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{dtree-3va-pruned.pdf}
  \caption{Cluster 3 vs all}
  \label{fig:3va}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{dtree-4va-pruned.pdf}
  \caption{Cluster 4 vs all}
  \label{fig:4va}
\end{figure}
\section{Other Work}
